Perturbing SU($n$)-symmetric interactions

### 2019-12-05 collective_sun/sun_perturbation.tex
fixed an error in the two-body perturbation theory

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\counts}{counts}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
  &&
  \V_2 \equiv \f12 \sum_{p,q} w_{pq} X_p Y_q
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $X,Y$ are trace-zero operators on an $n$-dimensional Hilbert
space; $X_p,Y_p$ denote the action of $X,Y$ on spin $p$; $\v X,\v Y$
are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
respectively a vector and matrix of all $v_p,w_{pq}$; we enforce
$w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to denote an
inner product between vectors and matrices as appropriate.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v \equiv \f1N \sum_p v_p,
  &&
  \col{X} \equiv \sum_p X_p.
\end{align}
In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  &= E_0 \V_1 \ket\psi
  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q v_q h_q X_q} \ket\psi \\
  &= E_0 \V_1 \ket\psi
  + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
  \label{eq:H_V1_psi}
\end{align}
where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of the net interaction energies $h_q\equiv\sum_ph_{pq}$ associated
with individual spins $p$; and $\diag\v h$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  We thus find that the vector
$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the coefficients $\v v$ satisfy
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v.
  \label{eq:cond_1}
\end{align}
If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
choice of single-spin trace-zero operator $X$ used to build $\V_1$ in
\eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
that generate eigenvectors of the interaction Hamiltonian $H_0$ when
they act on fully symmetric states $\ket\psi$ thus reduces to finding
eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \P_0 \V_2 \P_0
  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}},
\end{align}
where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
$B$; and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  \col{X\odot Y} \equiv \sum_p X_p\odot Y_p.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  &= E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
  &= E_0 \V_2 \ket\psi
  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
  - \v\1 \c \p{\m h\oc\m w} \c \v{X\odot Y} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
  &&
  h_k \equiv \sum_\ell h_{k\ell};
\end{align}
$\m w,\m h,\bar{\m h}$ are matrices of all
$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\m x\oc\m y$ is an element-wise
(Hadamard) product of $\m x$ and $\m y$, i.e.~with matrix elements
$x_{k\ell} y_{k\ell}$; $\v\1\equiv\p{1,1,1,\cdots}$ is an
$N$-component vector of ones; and $\v{X\odot Y}$ is a vector of all
$X_k\odot Y_k$.  We thus find that the vector $\V_2\ket\psi$ is an
eigenvector of the interaction Hamiltonian $H_0$ if the coefficient
matrix $\m w$ of all $w_{pq}$ satisfies
\begin{align}
  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
Remembering that the matrices $\m h$ and $\m w$ are both symmetric
(under transposition) with zeros on the diagonal, the diagonal
components of the condition in \eqref{eq:cond_2} automatically enforce
that the last term ($\sim\v{X\odot Y}$) in \eqref{eq:H_V2_psi}
vanishes.

We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
problem by writing the coefficient matrix $\m w$ as a vector
$\v{\m w}$ in $\RR^N\times\RR^N$,
\begin{align}
  \m w = \sum_{p,q} w_{pq} \op{p}{q}
  \to \v{\m w} \equiv \sum_{p,q} w_{pq} \ket{pq},
\end{align}
and defining the matrix
\begin{align}
  \tilde{\m h}
  \equiv \sum_{p,q} \bar h_{pq} \op{pq}
  + \sum_{p,q,\ell} h_{pq} \op{p\ell}{q\ell},
  \label{eq:h_super_mat}
\end{align}
in terms of which the condition in \eqref{eq:cond_2} becomes the
eigenvalue equation
\begin{align}
  \tilde{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
  \label{eq:cond_2_eig}
\end{align}
The matrix $\tilde{\m h}$ in \eqref{eq:h_super_mat} can be written in
the block-diagonal form
\begin{align}
  \tilde{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
  &&
  \m h_\ell
  \equiv \sum_{p,q}\p{\delta_{pq} \bar h_{p\ell} + h_{pq}} \op{p}{q},
\end{align}
which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
$N$ smaller, independent eigenvalue equations
\begin{align}
  \m h_\ell \c \v w_\ell = \f12 \Delta\p{\m w} \v w_\ell,
  &&
  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p},
  \label{eq:cond_2_eig_block}
\end{align}
where $\v w_\ell$ is essentially the $\ell$-th row or column of
$\m w$.  We can (numerically) diagonalize each of the blocks
$\m h_\ell$, and classify their eigenvectors $\v w_\ell$ by the
corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
$\v{\m w}$ of $\tilde{\m h}$ are then built from eigenvectors
$\v w_\ell$ within each block $\m h_\ell$ that share the same
eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
$\V_2\p{\m w_\Delta}\ket\psi$ of $H_0$ and determining their energy
thus reduces to finding the eigenvectors and eigenvalues of $N$
matrices $\m h_\ell$ of size $N\times N$.

If the interaction Hamiltonian $H_0$ is translationally invariant,
then $\bar h_{k\ell}=h_{k\ell}-h$, and $h_{k\ell}$ depends only on the
displacement between the spins indexed by $k$ and $\ell$.  The blocks
$\m h_\ell$ then take the form
\begin{align}
  \m h_{\ell} = \m\Sigma_\ell \c \m h_0 \c \m\Sigma_\ell^\dag,
  &&
  \m\Sigma_\ell \equiv \sum_p \op{p+\ell}{p},
  \label{eq:block_shift}
\end{align}
which implies that we need only find the eigenvectors and eigenvalues
of a single block $\m h_0$, after which we can determine the
eigenvectors of any block simply by appropriate applications of the
shift matrix $\m\Sigma_\ell$.  Note that in dimensions $D>1$, the
relation in \eqref{eq:block_shift} holds for vectors $p,\ell$ that
index lattice sites, i.e.~$p=\p{p_1,p_2,\cdots,p_D}$ and likewise with
$\ell$.

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue
$\Delta/2$, i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to
expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:PXYXYP} and \ref{sec:multi_to_collective}.  If
$X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
operator, then we can expand
\begin{align}
  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
neglected scalar terms in $H_2^{(2)}$ without physical consequence.
In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
$\m m$ we define the vectors
\begin{align}
  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
  &&
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
using which we have
\begin{align}
  W_4\p{\m h,\m w}
  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
\end{align}
\begin{align}
  W_2\p{\m h,\m w}
  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}
Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
the interaction matrix $\m h$ is implicit in the decomposition of the
coefficient matrix $\m w$ into components $\m w_\Delta$ as
$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.

@@

We wish to simplify
\begin{align}
  H_0 \V_2 \ket\psi
  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
  \Pi_{pq} X_k Y_\ell \ket\psi,
  \label{eq:H_V2_psi_start}
\end{align}
which has one term with $\set{p,q}=\set{k,\ell}$, terms with
$p,q\notin\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
permutation operator $\Pi_{pq}$ acts trivially on
$X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
\begin{align}
  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
  = \sum_{p<q} h_{pq}
  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  = E_0 + 2 h_{k\ell} - h_k - h_\ell
  = E_0 + 2 \bar h_{k\ell},
\end{align}
where
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
\end{align}
The corresponding terms in \eqref{eq:H_V2_psi_start} with
$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
\begin{align}
  E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
\end{align}
The terms in \eqref{eq:H_V2_psi_start} with only one of
$p,q\in\set{k,\ell}$ simplify to
\begin{align}
  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{k\ell}
  \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi
\end{align}
where $Z_k\equiv\p{X_kY_k+Y_kX_k}/2$.  In total,
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi.
\end{align}

@@

  $P\equiv\set{p}$ of $S$ consists of mutually disjoint, non-empty

@@

The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
consists of all operators that

@@

For any set $S$ of single-spin operators or partition $P$ of $S$, we
can similarly define the collective operators

@@

$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, all $S$), by use of a
computer algebra system (namely, Mathematica) we find that

@@

  \times \prod_{p\in\tilde P} \p{\abs{p}-1}!

@@

in $S$, and we enforce that partitions $\tilde P$ of an ordered list
$\Pi$, as well as parts $p\in\tilde P$, are lexicographically ordered
in accordance with the ordering of operators in $\Pi$.  If

@@

  O_{\tilde P} \equiv \bigotimes_{p\in\tilde P} \prod_{O\in p} O,

@@

  \equiv \prod_{p\in\tilde P} \col{\textstyle\prod_{O\in p} O},

@@

in which case we can also expand

>>>>

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\diag}{diag}
\let\flat\relax
\DeclareMathOperator{\flat}{flat}
\DeclareMathOperator{\counts}{counts}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
  &&
  \V_2 \equiv \sum_{p<q} w_{pq} \times \f12 \p{X_p Y_q + X_q Y_p}
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $X,Y$ are self-adjoint\footnote{In future versions of these
  notes, we should relax the restriction that $X,Y$ are self-adjoint.}
trace-zero operators on an $n$-dimensional Hilbert space; $X_p,Y_p$
denote the action of $X,Y$ on spin $p$; $\v X,$ $\v Y$, $\v v$, and
$\m w$ are respectively vectors and a matrix of all
$X_p,Y_p,v_p,w_{pq}$; we use a dot ($\c$) to denote an inner product
between vectors and matrices as appropriate; and we enforce
$w_{pq}=w_{qp}$ and $w_{pp}=0$.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v \equiv \f1N \sum_p v_p,
  &&
  \col{X} \equiv \sum_p X_p.
\end{align}
In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi,
  \label{eq:H_V1_psi}
\end{align}
where $h_q\equiv\sum_p h_{pq}$ is the net interaction energy
associated with spin $p$.  We thus find that the vector $\V_1\ket\psi$
is an eigenvector of the interaction Hamiltonian $H_0$ if the
coefficients $\v v$ satisfy
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v,
  \label{eq:cond_1}
\end{align}
where $\v h,\m h$ are respectively a vector and matrix of all
$h_p,h_{pq}$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
zeros everywhere else; and $\Delta\p{\v v}$ is some constant that may
depend on $\v v$.  If we {\it construct} a perturbation
$\V_1\p{\v v_\Delta}$ with coefficients $\v v_\Delta$ that satisfy
\eqref{eq:cond_1} with eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
choice of single-spin trace-zero operator $X$ used to build $\V_1$ in
\eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
that generate eigenvectors of the interaction Hamiltonian $H_0$ when
they act on fully symmetric states $\ket\psi$ thus reduces to finding
eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \P_0 \V_2 \P_0
  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}},
\end{align}
where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
$B$; and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  \col{X\odot Y} \equiv \sum_p X_p\odot Y_p.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
  X_k Y_\ell \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
  &&
  h_k \equiv \sum_\ell h_{k\ell}.
\end{align}
We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
interaction Hamiltonian $H_0$ if the coefficients $\m w$ satisfy
\begin{align}
  2\bar{\m h} \circ \m w + \m h\c\m w + \m w\c\m h
  = \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
where $\m w,\m h,\bar{\m h}$ are matrices of all
$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\bar{\m h} \circ \m w$ is an
element-wise (Hadamard) product of $\bar{\m h}$ and $\m w$, i.e.~with
matrix elements $\bar h_{k\ell} w_{k\ell}$; and $\Delta\p{\m w}$ is
some constant that may depend on $\m w$.

We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
problem by writing the coefficient matrix $\m w$ as a vector
$\v{\m w}$ in $\RR^N\times\RR^N$,
\begin{align}
  \m w = \sum_{p,q} w_{pq} \op{p}{q}
  \to \v{\m w} \equiv \flat\m w \equiv \sum_{p,q} w_{pq} \ket{pq},
\end{align}
where $\flat\m m$ is a flattening of the matrix $\m m$ into a vector
$\vec{\m m}$, and defining the matrix
\begin{align}
  \check{\m h}
  \equiv 2\diag\p{\flat\bar{\m h}}
  + \m h\otimes\1_N + \1_N\otimes\m h
  = 2 \sum_{p,q} \bar h_{pq} \op{pq}
  + \sum_{p,q,\ell} h_{pq} \p{\op{p\ell}{q\ell} + \op{\ell p}{\ell q}},
  \label{eq:h_super_mat}
\end{align}
where $\1_D$ is a $D\times D$ identity matrix.  The condition in
\eqref{eq:cond_2} is then equivalently the eigenvalue equation
\begin{align}
  \p{\check{\m d} \c \check{\m h} \c \check{\m d}} \c \v{\m w}
  = \Delta\p{\m w} \v{\m w},
  &&
  \check{\m d} \equiv \1_{N^2} - \sum_\ell \op{\ell\ell},
  \label{eq:cond_2_eig}
\end{align}
where $\check{\m d}$ annihilates the diagonal part of a (flattened)
matrix it acts on, while keeping the off-diagonal components the same.
We conjugate $\check{\m h}$ by $\check{\m d}$ to enforce that its
eigenvectors $\v{\m w}$ correspond to matrices $\m w$ with zeros on
the diagonal, i.e.~$w_{pp}=0$.

Both numerical experiments and physical arguments suggest that the
$N^2\times N^2$ matrix $\check{\m d} \c \check{\m h} \c \check{\m d}$
is highly degenerate, which means that we might be able to simplify
\eqref{eq:cond_2_eig} even further.  We defer this simplification to
future work, and for now bear the (large) computational cost of
diagonalizing a $N^2\times N^2$ matrix.

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue $\Delta$,
i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:PXYXYP} and \ref{sec:multi_to_collective}.  If
$X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
operator, then we can expand
\begin{align}
  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
neglected scalar terms in $H_2^{(2)}$ without physical consequence.
In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
$\m m$ we define the vectors
\begin{align}
  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
  &&
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
using which we have
\begin{align}
  W_4\p{\m h,\m w}
  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
\end{align}
\begin{align}
  W_2\p{\m h,\m w}
  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}
Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
the interaction matrix $\m h$ is implicit in the decomposition of the
coefficient matrix $\m w$ into components $\m w_\Delta$ as
$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
of $\m w$ onto the $\Delta$-eigenspace of the operator
$\check{\m d}\c\check{\m h}\c\check{\m d}$ defined by
\eqref{eq:h_super_mat} and \eqref{eq:cond_2_eig}.

@@

We wish to simplify
\begin{align}
  H_0 \V_2 \ket\psi
  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
  \Pi_{pq} X_k Y_\ell \ket\psi,
  \label{eq:H_V2_psi_start}
\end{align}
which has one term with $\set{p,q}=\set{k,\ell}$, terms with
$p,q\notin\set{k,\ell}$, and terms with only one of
$p,q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
trivially on $X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
\begin{align}
  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
  = \sum_{p<q} h_{pq}
  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  = E_0 + 2 h_{k\ell} - h_k - h_\ell
  = E_0 + 2 \bar h_{k\ell},
\end{align}
where
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
\end{align}
The corresponding terms in \eqref{eq:H_V2_psi_start} with
$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
\begin{align}
  E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
\end{align}
Using similar techniques, we can simplify the terms in
\eqref{eq:H_V2_psi_start} with only one of $p,q\in\set{k,\ell}$ to
\begin{align}
  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
  X_k Y_\ell \ket\psi.
\end{align}
In total,
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
  X_k Y_\ell \ket\psi.
\end{align}

@@

  $P\equiv\set{p}$ of $S$ is a set of mutually disjoint, non-empty

@@

In words, a partition $P$ of $S$ consists of an assignment of
operators $O\in S$ to individual parts $p\subset S$; each part $p$ is
associated with a tensor factor of $O_P$, and the operators $O\in p$
act on that tensor factor.  The set of multi-local operators
$\set{O_P:\t{partitions}~P~\t{of}~S}$ consists of all operators that

@@

For any set $S$ of single-spin operators and any partition $P$ of $S$,
we can similarly define the collective operators

@@

$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
a computer algebra system (namely, Mathematica) we find that

@@

  \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!

@@

in $S$; and we enforce that partitions $\tilde P$ of an ordered list
$\Pi$, as well as parts $\tilde p\in\tilde P$, are lexicographically
ordered in accordance with the ordering of operators in $\Pi$.  If

@@

  O_{\tilde P}
  \equiv \bigotimes_{\tilde p\in\tilde P} \prod_{O\in\tilde p} O,

@@

  \equiv \prod_{\tilde p\in\tilde P}
  \col{\textstyle\prod_{O\in\tilde p} O},

@@

where products over elements of $\tilde S,\tilde P,\tilde p$ are taken
in order according to the elements of these ordered lists, e.g.~if
$\tilde S=\p{O_1,O_2,\cdots,O_{\abs{S}}}$, then
$\prod_{O\in\tilde S}\equiv O_1O_2\cdots O_{\abs{S}}$.  These
definitions allow us to expand
